---
title: "FSIM_spatial_analysis"
author: "Claire Tortorelli"
date: "12/20/2020"
output: html_document
---

Analyses for BME BP and FIL > 4ft
```{r setup, include=FALSE}

library(tidyverse)
library(here)
library(ggridges)
library(cowplot)
library(ggridges)
```

read in RData
```{r}
load(here("saptialcsv_rawdata", "AllData_wFIL_120m_point_extract_cleaned.RData")) 
#dfsub
```


```{r}
# summary(dfsub)
# 
# #add column for BP * FIL > 4ft
# dfsub$foa10FILBP <- dfsub$foa10FILg4 * dfsub$foa10BP
# dfsub$foa12FILBP <- dfsub$foa12FILg4 * dfsub$foa12BP
# dfsub$difFILBP <- dfsub$foa10FILBP - dfsub$foa12FILBP

```

```{r}
#calcualte area burned at each flame length
studyArea <- nrow(dfsub)*1.44 #each cell represents 1.44 ha

sum(dfsub$foa10FILg4*1.44) #3697106, 54.9% of ecoregion
sum(dfsub$foa12FILg4*1.44) #3622070, 53.8% of ecoregion

sum(dfsub$foa10FIL8*1.44) #748102.2, 11.1%
sum(dfsub$foa12FIL8*1.44) #737993, 11.0%

#percent difference
(sum(dfsub$foa10FILg4*1.44)-sum(dfsub$foa12FILg4*1.44))/sum(dfsub$foa12FILg4*1.44)*100 #2.1
(sum(dfsub$foa10FIL8*1.44)-sum(dfsub$foa12FIL8*1.44))/sum(dfsub$foa12FIL8*1.44) *100 #1.4

#difference in BP
mean(dfsub$foa10BP) - mean(dfsub$foa12BP)
```

summarize burn probability just in invaded areas
```{r}
#create a df with only invaded areas where we increased fuels between landscapes
invaded.df <- dfsub[which(dfsub$alteredFM2 == 1),]
invaded.df$reclass <- "Invaded"

invaded.df.sum <- invaded.df %>%
    summarize(mean_F4dif = mean(difFILg4), #calculate absolute difference FIL > 4gt
              mean_BPdif = mean(difBP), #calculate absolute difference BP
              # mean_FLBPdif = mean(dfsub$difFILBP), #calculate absolute difference BP*FL
              mean_F8dif = mean(difFIL8),
              F4_percdif = ((mean(foa10FILg4) - mean(foa12FILg4)) / (mean(foa12FILg4)))*100,
              F8_percdif = ((mean(foa10FIL8) - mean(foa12FIL8)) / (mean(foa12FIL8)))*100,
              BP_percdif = ((mean(foa10BP) - mean(foa12BP)) / (mean(foa12BP)))*100,
              # FLBP_percdif = ((mean(dfsub$foa10FILBP) - mean(dfsub$foa12FILBP)) / (mean(dfsub$foa12FILBP)))*100,
              n = n() #count
              )  

#summary(invaded.df)
nrow(invaded.df)/nrow(dfsub)
```


summarize mean BP and MI by EVT (vegetation type) for the ecoregion - ignore NAs for calculating percent difference
```{r}

sum.EVTall <- dfsub %>%
  group_by(reclass) %>%
    summarize(mean_F410 = mean(foa10FILg4),
              mean_F412 = mean(foa12FILg4),
              mean_BP10 = mean(foa10BP),
              mean_BP12 = mean(foa12BP),
              mean_F810 = mean(foa10FIL8),
              mean_F812 = mean(foa12FIL8),
              mean_F4dif = mean(difFILg4), #calculate absolute difference FIL > 4ft
              mean_BPdif = mean(difBP), #calculate absolute difference BP
              mean_F8dif = mean(difFIL8),#calculate absolute difference FIL > 8ft  
              F4_percdif = ((mean(foa10FILg4) - mean(foa12FILg4)) / (mean(foa12FILg4)))*100,
              BP_percdif = ((mean(foa10BP) - mean(foa12BP)) / (mean(foa12BP)))*100,
              F8_percdif = ((mean(foa10FIL8) - mean(foa12FIL8)) / (mean(foa12FIL8)))*100,
              n = n() #count
              )  

#write.csv(sum.EVTall, "meanBP.MI.FH.by.EVT.for.all.burnable.areas.csv")
```

```{r}
#explore why sparse veg values are so high?
# sparsevegn <- dfsub %>% subset(reclass == "Sparsely vegetated") %>%
#   group_by(PRE_FM40) %>%
#   summarize(n = n())
#looks like LANDFIRE models most of these areas as FM122, GR2! 

# recentydistn <- dfsub %>% subset(reclass == "Recently Disturbed") %>%
#   group_by(PRE_FM40) %>%
#   summarize(n = n())

#  

```

There's hardly any differnce in area burned at >8ft so sticking with 4ft

subset and gather dataframes to plot invaded (foa10) and uninvaded (foa12) runs together for comparison
```{r}

#FIL>4 - extract OID, reclass, and vars (reclassified vegtype)
dfsubl_F4 <- dfsub[,c(2,11,49,48)]
dfsubl_F4 <- dfsubl_F4 %>% gather("FSIMrun", "F4", foa10FILg4, foa12FILg4)

#BP 
dfsubl_BP <- dfsub[,c(2,5,11,17)]
dfsubl_BP <- dfsubl_BP %>% gather("FSIMrun", "BP", foa10BP, foa12BP)


#FILBP
dfsubl_FLBP <- dfsub[,c(2,11,51,52)]
dfsubl_FLBP <- dfsubl_FLBP %>% gather("FSIMrun", "FLBP", foa12FILBP, foa10FILBP)



```





organize means for plotting with the density graphs
```{r}

densitymeansF4 <- dfsubl_F4 %>% group_by(reclass, FSIMrun) %>%
  summarise(mean = mean(F4))

densitymeansBP <- dfsubl_BP %>% group_by(reclass, FSIMrun) %>%
  summarise(mean = mean(BP))

densitymeansFLBP <- dfsubl_FLBP %>% group_by(reclass, FSIMrun) %>%
  summarise(mean = mean(FLBP))



```


### Plot
plot MI densities for all veg types for invaded and uninvaded landscapes

Function to plot burn metrics vs vegtype in ggplot
```{r fig.height=7, fig.width=6}
vegcolors <- c("#086375", "#b2b2b2", "#A06C18", "#883C0C", "#598B2C", "#d7ce91", "#ffd500", "#386641", "#a6c952")
fsimcolors <- c("#cc5500", "#323232")
fsimcolors2 <- c("#cc4125", "#323232")

ggp_function <- function(df, x.var, df2){
  ggp <- ggplot(df, aes(x = x.var, y = reclass, color = FSIMrun, fill = reclass)) +

  geom_density_ridges(rel_min_height = 0.01, alpha = 0.3, quantile_lines = TRUE, quantiles = 2, scale = 1.1)+
        theme(plot.title = element_text(size = 12)) +
  scale_fill_manual(values = vegcolors)+ 
    scale_alpha_manual(values = c(0.1, 0.5))+
  theme_bw(base_size = 13)+
  theme(panel.grid.minor = element_blank
        (), panel.grid.major = element_blank(),
        axis.title.y = element_blank(),
       legend.position = "none")
  
  gp_point <- ggp + geom_point(aes(x = mean, #add means as points to each density plot
                   y = reclass,
                   color = FSIMrun),
                   size = 2.5,
                   shape = 17,
                   data = df2)+
       scale_color_manual(values = fsimcolors)
  
  return(gp_point)
  
  }


```

Reverse order of vegtypes for plotting
```{r}
dfsubl_F4$reclass <- fct_rev(factor(dfsubl_F4$reclass))
dfsubl_BP$reclass <- fct_rev(factor(dfsubl_BP$reclass))
dfsubl_FLBP$reclass <- fct_rev(factor(dfsubl_FLBP$reclass))
levels(dfsubl_BP$reclass)
```

plot for burn prob, mean intensity, and fire hazard
```{r fig.height=7, fig.width=6}


p_bp <- ggp_function(df = dfsubl_BP, x.var = dfsubl_BP$BP, df2 = densitymeansBP) + xlab('Burn prob.') + ggtitle("Annual burn prob.") +  scale_x_continuous(limits = c(0, 0.035))


p_f4 <- ggp_function(df = dfsubl_F4, x.var = dfsubl_F4$F4, df2 = densitymeansF4) + xlab('Flame-length prob. >1.2m') + ggtitle("Flame-length prob. >1.2m")


# 
# p_flbp <- ggp_function(df = dfsubl_FLBP, x.var = dfsubl_FLBP$FLBP, df2 = densitymeansFLBP) + xlab('Joint prob.') + ggtitle("Joint prob.")+  theme(axis.title.y=element_blank(),
#         axis.text.y=element_blank())+  scale_x_continuous(limits = c(0, 0.026))
# 

```

```{r}
#library(egg)
P_all <- ggarrange(p_bp, 
          p_f4 + 
               theme(axis.text.y = element_blank(),
                     #axis.ticks.y = element_blank(),
                     axis.title.y = element_blank() ),
                nrow = 1)
```


save plot

```{r}
#save plot
svg(file = 'density_plots_by_vegTypesFL.svg', width = 7, height = 6.5)
P_all
dev.off()
```


