---
title: "FSIM_spatial_analysis"
author: "Claire Tortorelli"
date: "12/20/2020"
output: html_document
---

# Spatial patterns of Fire: Patch and landscape scripts


### Set up
```{r setup, include=FALSE}
library(tidyverse)
library(raster)
library(landscapemetrics)
library(here)
library(ggplot2)
library(cowplot)
library(grid)
library(gridExtra) 
library(gam)
```



read in RData
```{r}
load(here("spatialcsv_rawdata", "AllData_cleaned_2022.RData"))
# dfsub = read.csv(here('spatialcsv_rawdata', 'AllData_120m_point_extract.csv'))
#dfsub
```



Covert to spatial data
```{r}
# #convert to spatial points
coords <- data.frame(dfsub[,3:4])
sp = SpatialPoints(coords) 

#set projection: NAD_1983_USFS_R6_Albers;  coordinate system:GCS_North_American_1983  EPSG:5070 - for entire US

spdf = SpatialPointsDataFrame(sp, dfsub)

```





Create function to create rasters from different spatial points for each burn metric
```{r}
#convert spatial df to raster

#function to create raster from coordinates
coords2raster <- function(z.input, xy){ 
      #absdif = absolute difference fire metric vector from dfsu; 
      #xy = are x and y coordinates from entire study area

      #set raster value to difference in fire behavior metric
      xy$Z <- z.input
  
      #convert to raster from coordinates
      raster_output <- rasterFromXYZ(xy)
}
```

create rasters for burn metrics (absolute difference)
```{r}

BPraster <- coords2raster(dfsub$difBP, coords) #difference in BP
#plot(BPraster)

F4raster <- coords2raster(dfsub$difFILg4, coords) #differnece in bp>1.2m (4ft)


F8raster <- coords2raster(dfsub$difFIL8, coords) #BP 2.4m raster (8ft)
```

create rasters for burn metrics (absolute difference)
```{r}

BP_10raster <- coords2raster(dfsub$foa10BP, coords) # BP invaded run
BP_12raster <- coords2raster(dfsub$foa12BP, coords) #BP uninvaded run

F4_10raster <- coords2raster(dfsub$foa10FILg4, coords) #BP>1.2m invaded run
F4_12raster <- coords2raster(dfsub$foa12FILg4, coords) #BP>1.2m uninvaded run


F8_12raster <- coords2raster(dfsub$foa12FIL8, coords) #BP>2.4m uninvaded run
F8_10raster <- coords2raster(dfsub$foa10FIL8, coords) #BP>2.4m invaded run
```

## Patch analysis

create raster for invaded/ fuel altered status (1 = fuels were altered to reflect invasion)
```{r}
#add zeros to indicate univaded areas
dfsub$alteredFM2[is.na(dfsub$alteredFM2)] <- 0

#extract invaded status into new raster
invraster_BME <- coords2raster(dfsub$alteredFM2, coords)
#plot(invraster_BME)
```


Extract invaded areas to separate raster for patch size calculations
```{r}
#extract coordinates for just invaded areas
spdf_invaded <- dfsub[which(dfsub$alteredFM2 == 1),]
coords_invaded <- spdf_invaded[,3:4]

#convert to spatial points
sp_invaded = SpatialPoints(coords_invaded)

invraster <- coords2raster(spdf_invaded$alteredFM2, coords_invaded)
plot(invraster)

```

### Calcualte patch area
```{r}
library(landscapemetrics)
#calculate area of each invaded patch in hectares, Queens case
#extract patch area for invaded patches 
pts_extract <- extract_lsm(invraster, y=spdf, what ="lsm_p_area", directions = 8) 

# area values are reported in hectares

pts_extract$area <- pts_extract$value #assign patch area column 

```


extract absolute difference fire metric raster values to invaded spatial points 
```{r}
#extract difference in burn probability to invaded patches with area measurements
pts_extract$bp_dif <- raster::extract(BPraster, sp_invaded)
pts_extract$F4_dif <- raster::extract(F4raster, sp_invaded)

#RawBP& FIL4
pts_extract$foa10BP <- raster::extract(BP_10raster, sp_invaded) 
pts_extract$foa12BP <- raster::extract(BP_12raster, sp_invaded) 
pts_extract$foa10F4 <- raster::extract(F4_10raster, sp_invaded) 
pts_extract$foa12F4 <- raster::extract(F4_12raster, sp_invaded) 
```

calculate mean burn metric for each patch
```{r}
#average fire metrics by patch ids
pts_extractdf <- data.frame(pts_extract)

patch_means <- pts_extractdf %>% 
  group_by(id) %>%
  summarise(meanbp_dif = mean(bp_dif),
            meanF4_dif = mean(F4_dif),
            meanBP_10 = mean(foa10BP),
            meanBP_12 = mean(foa12BP),
            meanF4_10 = mean(foa10F4),
            meanF4_12 = mean(foa12F4),
            area = mean(area))

summary(patch_means)
```

### Model burn metric ~ patch size with GAM

Absolute differences in patch metrics - these are modeled with Gaussian distribution because they have some negative numbers and don't approach 1
```{r}
#assign weights to patch_means for binomial models
patch_means$trials = 10000

mod_gamBPpatch <- gam::gam(meanbp_dif ~ s(log(area)), data=patch_means)
#mod_lmBP <- lm(meanbp_dif ~ log(area), data=patch_means)
#no sig difference between lm and gam, but gam appears to fit the data better at higher patch sizes
mod_gamF4patch <- gam::gam(meanF4_dif ~ s(log(area)), data=patch_means)

#summary
summary(mod_gamBPpatch)
summary(mod_gamF4patch)

#predict to new data
newdat <- data.frame(area = c(1.44, 2.88, 12, 150, 8650))
predict(mod_gamBPpatch, newdat, type = "response")
predict(mod_gamF4patch, newdat, type = "response")
#BP range between 0.0009660715 and 0.0037170969 
#F4range between 0.19 and *0.31* for the smallest and largest patches
```

model burn metric response by simulation
```{r}
mod_gamBP10patch <- gam::gam(meanBP_10 ~ s(log(area)), family = "binomial", data = patch_means, weights = trials)
mod_gamBP12patch <- gam::gam(meanBP_12 ~ s(log(area)), family = "binomial", data = patch_means, weights = trials)
mod_gamF410patch <- gam::gam(round(meanF4_10, digits = 6) ~ s(log(area)), family = "binomial", data = patch_means, weights = trials)
mod_gamF412patch <- gam::gam(round(meanF4_12, digits = 6) ~ s(log(area)), family = "binomial", data = patch_means, weights = trials)

#summary
summary(mod_gamBP10patch)
summary(mod_gamBP12patch)
summary(mod_gamF410patch)
summary(mod_gamF412patch)

#predict
newdat
predict.Gam(mod_gamBP10patch, newdat, type = "response")
predict.Gam(mod_gamBP12patch, newdat, type = "response")
(0.007680588 - 0.006233901)/ 0.006233901 
#predict to new data for plotting

df_new <- data.frame(area = seq(1,8650, length.out = 1000))

df_new$meanBP_10 <- predict.Gam(mod_gamBP10patch, newdata = df_new, type = "response")
df_new$meanBP_12 <- predict.Gam(mod_gamBP12patch, newdata = df_new, type = "response")
df_new$meanF4_10 <- predict.Gam(mod_gamF410patch, newdata = df_new, type = "response")
df_new$meanF4_12 <- predict.Gam(mod_gamF412patch, newdata = df_new, type = "response")
```

### plot patch relationships with changes in burn metrics
```{r plot patch}
#plot BP
(ggbp_patch <- ggplot(data = df_new, aes(x = log(area)))+
  #geom_jitter(alpha = .5, size = 2)+
    geom_smooth(color = "#cc5500", aes(y = meanBP_10), se = FALSE) +
    geom_smooth(color = "#323232", linetype = "dashed", aes(y = meanBP_12), se = FALSE) +
  theme_bw(base_size = 13)+
  #scale_color_manual(values = colors)+ #colors for vegtypes
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  xlab("")+
  ylab("BP")+
  theme(plot.title = element_text(hjust = 0.5))
)



#plot flame lengths
(ggflep_patch <- ggplot(data = df_new, aes(x = log(area)))+
  #geom_jitter(alpha = .5, size = 2)+
    geom_smooth(color = "#cc5500", aes(y = meanF4_10), se = FALSE) +
    geom_smooth(color = "#323232", linetype = "dashed", aes(y = meanF4_12), se = FALSE) +
  theme_bw(base_size = 13)+
  #scale_color_manual(values = colors)+ #colors for vegtypes
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  labs(y=expression("BP"[">1.2m"]), x = "")+
  theme(plot.title = element_text(hjust = 0.5))
)


```


Combine plots
```{r fig.height=4, fig.width=6}

#create common x axis label
x.grob <- textGrob("invaded patch size (log ha)", vjust = -0.5)

#combine plots
p_all <- plot_grid(ggbp_patch, ggflep_patch  + theme(legend.position = "none"),
                    ncol = 2,
                    nrow = 1,
                    labels = c('(a)', '(b)'),
                    label_x = 0, label_y = 0,
                    hjust  = -0.5, vjust = -2.5)

p <- plot_grid(p_all, nrow = 2, rel_heights = c(1, .1))
```

save plot
```{r}
svg(file = 'invaded_patchsizeGAM_all_binom.svg', width = 6, height = 3.5)

#add common x label
grid.arrange(arrangeGrob(p_all, bottom = x.grob))

dev.off()
```


-----------------------------------------------------------------------------------

## Focal Neighborhood analysis


Calculate neighborhood metrics for all invaded pixels - this is the amount of invasion within 116.6 hectares. We chose this neighborhood size to reflect "large" fire sizes included in the transition analysis. 
```{r}

focal <- raster::focal(invraster_BME, w=matrix(1/9^2,nrow=9,ncol=9), na.rm = TRUE) #sum of focal neighborhood in number of cells - a 5x5 cell kernel–centered on the cell whose value is being estimated. - this needs to be an odd number of cells

#120m x 5 cells = 600m row and 600m cols = 360,000 m^2 = 36 hectares
#120m x 7 cells = 840m row and 840m cols = 705,600 m^2 = 70.56 hectares
#120m x 9 cells = 1080m row and 1080m cols = 1,166,400m^2 = 116.64 hectares


plot(focal) #plots neighborhood map. 
# plot(invraster_BME)


#summary(focal)
```

Examine relationship between difference in burn probability and amount of invaded cells within neighborhood

extract values from focal raster to spatial dataframe
```{r}
#extract difference in burn probability to invaded patches with area measurements
spdf$focal_inv <- raster::extract(focal, sp) 

#subset so only areas with > 0 invaded area within  neighborhood are included in analysis/plot
spdf_sub <- data.frame(spdf[which(spdf$focal_inv > 0),])



#keep only forested areas
spdf_forest <- data.frame(spdf_sub[which(spdf_sub$reclass == "Open tree canopy" | spdf_sub$reclass == "Closed tree canopy"),])

summary(spdf_sub$alteredFM2)

#remove invaded focal cells from analysis
#spdf_forestuninv <- data.frame(spdf_forest[which(spdf_forest$alteredFM2 == 0),])
spdf_forestuninv <- data.frame(spdf_forest[is.na(spdf_forest$alteredFM2),])

```

### model relationships using GAM


Absolute dif. for burn metrics
```{r}
mod_gamBP <- gam::gam(difBP ~ s(focal_inv), data=spdf_forestuninv)
mod_gamF4 <- gam::gam(difFILg4 ~ s(focal_inv), data=spdf_forestuninv)
mod_gamF8 <- gam::gam(difFIL8 ~ s(focal_inv), data=spdf_forestuninv)

#summary
summary(mod_gamBP)
summary(mod_gamF4)
summary(mod_gamF8)

#predict
newdat2 <- data.frame(focal_inv = c(0, .25, .50, .100))

predict.Gam(mod_gamBP, newdat2) #0, 25%, 50%, 75% invaded
predict.Gam(mod_gamF4, newdat2)
predict.Gam(mod_gamF8, newdat2)

```



Predict values of BP and FLEP at differnet prop. invaded levels
```{r predict focal}
#add weights for binomial models
spdf_forestuninv$trials = 10000
#bp
mod_gambpfoa10 <- gam::gam(foa10BP ~ s(focal_inv), data=spdf_forestuninv,  family = "binomial")
mod_gambpfoa12 <- gam::gam(foa12BP ~ s(focal_inv), data=spdf_forestuninv, weights= trials,  family = "binomial")
#f4
mod_gamF4foa10 <- gam::gam(round(foa10FILg4, digits = 6) ~ s(focal_inv), data=spdf_forestuninv, weights= trials,  family = "binomial")
mod_gamF4foa12 <- gam::gam(round(foa12FILg4, digits = 6) ~ s(focal_inv), data=spdf_forestuninv, weights= trials,  family = "binomial")
#f8
mod_gamF8foa10 <- gam::gam(round(foa10FIL8, digits = 6) ~ s(focal_inv), data=spdf_forestuninv, weights= trials,  family = "binomial")
mod_gamF8foa12 <- gam::gam(round(foa12FIL8, digits = 6) ~ s(focal_inv), data=spdf_forestuninv, weights= trials,  family = "binomial")

#summary
summary(mod_gambpfoa10)
summary(mod_gambpfoa12)
summary(mod_gamF4foa10)
summary(mod_gamF4foa12)
summary(mod_gamF8foa10)
summary(mod_gamF8foa12)

#predict
preds3 <- data.frame(focal_inv = c(.10, .25, .50, .99))
predict.Gam(mod_gambpfoa10, preds3, type = 'response')
predict.Gam(mod_gambpfoa12, preds3, type = 'response')

#10% invaded
(0.008288111 - 0.007215339)/0.007215339 #15%
#25% invaded
(0.008158160    - 0.006321575 )/0.006321575   #0.0018;  28.3% increase *binom = 29.1%*
#50% invaded
(0.007761675    - 0.004921681 )/ 0.004921681    #0.0028; 57.9% increase *binom = 57.7%*
#99 % invaded
(0.008227249   - 0.003575630)/0.003575630 #128% *binom = 130%*

#repeat for BP >1.2m
predict.Gam(mod_gamF4foa10, newdat2, type = "response")
predict.Gam(mod_gamF4foa12, newdat2, type = "response")

#25% invaded
(0.5601069   - 0.5148772   )/0.5148772    #0.045; 8.7% *binom = 8.8%*
#50% invaded
(0.6008589     - 0.5085612  )/ 0.5085612    #0.0910; 17.9% *binom = 18.1%*

#repeat for BP >2.4m
predict.Gam(mod_gamF8foa10, newdat2, type = "response")
predict.Gam(mod_gamF8foa12, newdat2, type = "response")

#25% invaded
(0.09341769   - 0.08002696  )/0.08002696   #0.0136; 16.4% *binom = 16.7%*
#50% invaded 
(0.09435233   - 0.06510738   )/ 0.06510738   #0.0287; 42.7% *binom = 44.9%*

##predict to new data for plotting
df_new_neighborhood <- data.frame(focal_inv = seq(0.01,0.97, length.out = 1000))

df_new_neighborhood$foa10BP <- predict.Gam(mod_gambpfoa10, newdata = df_new_neighborhood, type = "response")
df_new_neighborhood$foa12BP <- predict.Gam(mod_gambpfoa12, newdata = df_new_neighborhood, type = "response")
df_new_neighborhood$foa10FILg4 <- predict.Gam(mod_gamF4foa10, newdata = df_new_neighborhood, type = "response")
df_new_neighborhood$foa12FILg4 <- predict.Gam(mod_gamF4foa12, newdata = df_new_neighborhood, type = "response")
df_new_neighborhood$foa10FIL8 <- predict.Gam(mod_gamF8foa10, newdata = df_new_neighborhood, type = "response")
df_new_neighborhood$foa12FIL8 <- predict.Gam(mod_gamF8foa12, newdata = df_new_neighborhood, type = "response")
```



### Plot focal neighborhood for forested areas

```{r}
#plot raw BP and FLEP1.2m values for the focal forested cells

# plot BP
(p_forestBP <- ggplot(data = df_new_neighborhood, aes(x = focal_inv, y = foa10BP))+
  #geom_jitter(alpha = .5, size = 2)+
    geom_smooth(color = "#cc5500", se = FALSE) +
    geom_smooth(color = "#323232", linetype = "dashed", aes(y = foa12BP), se = FALSE) +
  theme_bw(base_size = 13)+
  #scale_color_manual(values = colors)+ #colors for vegtypes
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  xlab("")+
  ylab("BP")+
  theme(plot.title = element_text(hjust = 0.5))
)

#plot F4
(p_forestF4 <- ggplot(data = df_new_neighborhood, aes(x = focal_inv, y = foa10FILg4))+
  #geom_jitter(alpha = .5, size = 2)+
     geom_smooth(color = "#cc5500", se = FALSE) +
    geom_smooth(color = "#323232", linetype = "dashed", aes(y = foa12FILg4), se = FALSE) +
  theme_bw(base_size = 13)+
  #scale_color_manual(values = colors)+ #colors for vegtypes
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  labs(y=expression("BP"[">1.2m"]), x = "")+
  theme(plot.title = element_text(hjust = 0.5))
)

#plot F8
(p_forestF8 <- ggplot(data = df_new_neighborhood, aes(x = focal_inv, y = foa10FIL8))+
  #geom_jitter(alpha = .5, size = 2)+
     geom_smooth(color = "#cc5500", se = FALSE) +
    geom_smooth(color = "#323232", linetype = "dashed", aes(y = foa12FIL8), se = FALSE) +
  theme_bw(base_size = 13)+
  #scale_color_manual(values = colors)+ #colors for vegtypes
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  labs(y=expression("BP"[">2.4m"]), x = "")+
  theme(plot.title = element_text(hjust = 0.5))
)

```





Combine plots
```{r fig.height=4, fig.width=8}

#create common x axis label
x.grob2 <- textGrob("Proportion of forest neighborhood invaded", vjust = -1)


#combine plots
p_allforest <- plot_grid(p_forestBP, p_forestF4, p_forestF8 + 
                           theme(legend.position = "none"),
                    ncol = 3,
                    nrow = 1,
                    labels = c('(a)', '(b)', '(c)'),
                    label_x = 0, label_y = 0,
                    hjust  = -0.5, vjust = -2.5)

p <- plot_grid(p_allforest, nrow = 1, rel_heights = c(1, .2))
```

save plot
```{r}
svg(file = 'focal_neighborhood116haGAM_all_binom.svg', width = 7.8, height = 2.7)

#add common x label
grid.arrange(arrangeGrob(p, bottom = x.grob2))

dev.off()
```



-------------------------------------------------------------------------------------

###  Calculate landscape fire change ~ propotion of invaded landscape


Average landscape burn metrics for 9x9 pixel landscape ~1000x1000m

```{r}

landscape_BP <- raster::focal(BPraster, fun = "mean", w=matrix(1,9,9), na.rm = TRUE) #focal areas with NA in the neighborhood have mean values as NA
landscape_F4 <- raster::focal(F4raster, fun = "mean", w=matrix(1,9,9), na.rm = TRUE)
landscape_F8 <- raster::focal(F8raster, fun = "mean", w=matrix(1,9,9), na.rm = TRUE)


```

plot relationship between difference in neighborhood burn metrics  and amount of invaded cells within 116.6ha

extract values from focal raster to spatial dataframe
```{r}
#extract difference in neighborhood burn metrics to the spatial dataframe
spdf$landscape_BP <- raster::extract(landscape_BP, sp) 
spdf$landscape_F4 <- raster::extract(landscape_F4, sp) 
spdf$landscape_F8 <- raster::extract(landscape_F8, sp) 


#subset so only areas with > 0 invaded area within neighborhood are included in plot
spdf_sub <- data.frame(spdf[which(spdf$focal_inv > 0),])
#summary(spdf_sub)

```


### Examine landscape neighborhood change


Average landscape burn metrics for 9x9 pixel landscape ~1000x1000m
```{r}

landscape_BP10 <- raster::focal(BP_10raster, fun = "mean", w=matrix(1,9,9), na.rm = TRUE) #focal areas with NA in the landscape have mean values as NA
landscape_BP12 <- raster::focal(BP_12raster, fun = "mean", w=matrix(1,9,9), na.rm = TRUE) 

#repeat for BP>1.2m
landscape_f410 <- raster::focal(F4_10raster, fun = "mean", w=matrix(1,9,9), na.rm = TRUE) 
landscape_f412 <- raster::focal(F4_12raster, fun = "mean", w=matrix(1,9,9), na.rm = TRUE) 

#repeat for BP>2.4m
landscape_f810 <- raster::focal(F8_10raster, fun = "mean", w=matrix(1,9,9), na.rm = TRUE) 
landscape_f812 <- raster::focal(F8_12raster, fun = "mean", w=matrix(1,9,9), na.rm = TRUE) 

```

extract values from focal raster to spatial dataframe
```{r}
#extract difference in landscape burn metrics to the spatial dataframe
spdf$landscape_BP10 <- raster::extract(landscape_BP10, sp) 
spdf$landscape_BP12 <- raster::extract(landscape_BP12, sp) 
spdf$landscape_f410 <- raster::extract(landscape_f410, sp) 
spdf$landscape_f412 <- raster::extract(landscape_f412, sp) 

#for BP>2.4
spdf$landscape_f810 <- raster::extract(landscape_f810, sp) 
spdf$landscape_f812 <- raster::extract(landscape_f812, sp) 



#subset so only areas with > 0 invaded area within neighborhood are included in plot
spdf_sub <- data.frame(spdf[which(spdf$focal_inv > 0),])
#summary(spdf_sub)
```

### model relationships with GAMs

```{r}
#difference plots
mod_gamBP_landscape <- gam::gam(landscape_BP ~ s(focal_inv), data=spdf_sub)
mod_gamF4_landscape <- gam::gam(landscape_F4 ~ s(focal_inv), data=spdf_sub)
mod_gamF8_landscape <- gam::gam(landscape_F8 ~ s(focal_inv), data=spdf_sub)

#summary
summary(mod_gamBP_landscape)
summary(mod_gamF4_landscape)
summary(mod_gamF8_landscape)

```

Model burn metric responses by simulation
```{r}
#model by simulation
spdf_sub$trials = 1000 #add weights for binomial gams

mod_gambpfoa10_landscape <- gam::gam(landscape_BP10 ~ s(focal_inv), data=spdf_sub, family = "binomial", weights = trials)
mod_gambpfoa12_landscape <- gam::gam(landscape_BP12 ~ s(focal_inv), data=spdf_sub, family = "binomial", weights = trials)

mod_gamF4foa10_landscape <- gam::gam(round(landscape_f410, 6) ~ s(focal_inv), data=spdf_sub, family = "binomial", weights = trials)
mod_gamF4foa12_landscape <- gam::gam(round(landscape_f412, 6) ~ s(focal_inv), data=spdf_sub, family = "binomial", weights = trials)
mod_gamF8foa10_landscape <- gam::gam(round(landscape_f810, 6) ~ s(focal_inv), data=spdf_sub, family = "binomial", weights = trials)
mod_gamF8foa12_landscape <- gam::gam(round(landscape_f812, 6) ~ s(focal_inv), data=spdf_sub, family = "binomial", weights = trials)


#predict #0, .25, .5, .75 prop landscape invaded
#BP
predict.Gam(mod_gambpfoa10_landscape, newdat2, type = "response")
predict.Gam(mod_gambpfoa12_landscape, newdat2, type = "response")

#percent increase for 25% invaded
(0.007468408  - 0.005771739 ) /0.005771739  #0.00167; 28.9% - binom = 29.4%
#percent increase for 50% invaded
(0.007897086  -0.005111080  )/0.005111080  #0.0027; 54% - binom = 54.5%

#BP>1.2m
predict.Gam(mod_gamF4foa10_landscape, newdat2, type = "response")
predict.Gam(mod_gamF4foa12_landscape, newdat2, type = "response")

#percent increase for 25% invaded
(0.5575173  -0.4595866  ) / 0.4595866  #0.098; 21% - binom same
#percent increase for 50% invaded
(.6295171  -0.4456754  )/0.4456754  # 0.184; 41% - same

#BP(>2.4m)
predict.Gam(mod_gamF8foa10_landscape, newdat2, type = "response")
predict.Gam(mod_gamF8foa12_landscape, newdat2, type = "response")


#percent increase for 25% invaded
(0.06115281 -0.05231815   ) / 0.05231815  #0.0088; 16% - binom 16.8%
#percent increase for 50% invaded
(0.06618273   -0.04460802  )/0.04460802  # 0.021; 47.7% - 48.4%


##predict to new data for plotting
df_new_landscape <- data.frame(focal_inv = seq(0.01,0.97, length.out = 1000))

df_new_landscape$landscape_BP10 <- predict.Gam(mod_gambpfoa10_landscape, newdata = df_new_landscape, type = "response")
df_new_landscape$landscape_BP12 <- predict.Gam(mod_gambpfoa12_landscape, newdata = df_new_landscape, type = "response")
df_new_landscape$landscape_f410 <- predict.Gam(mod_gamF4foa10_landscape, newdata = df_new_landscape, type = "response")
df_new_landscape$landscape_f412 <- predict.Gam(mod_gamF4foa12_landscape, newdata = df_new_landscape, type = "response")
df_new_landscape$landscape_f810 <- predict.Gam(mod_gamF8foa10_landscape, newdata = df_new_landscape, type = "response")
df_new_landscape$landscape_f812 <- predict.Gam(mod_gamF8foa12_landscape, newdata = df_new_landscape, type = "response")
```



```{r}


(p_landscapeBP <- ggplot(data = df_new_landscape, aes(x = focal_inv, y = landscape_BP10))+
  #geom_jitter(alpha = .5, size = 2)+
     geom_smooth(color = "#cc5500", se = FALSE) +
    geom_smooth(color = "#323232", linetype = "dashed", aes(y = landscape_BP12), se = FALSE) +
  theme_bw(base_size = 13)+
  #scale_color_manual(values = colors)+ #colors for vegtypes
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  xlab("")+
  ylab("BP")+
  theme(plot.title = element_text(hjust = 0.5))
)

(p_landscapeF4 <- ggplot(data = df_new_landscape, aes(x = focal_inv, y = landscape_f410))+
  #geom_jitter(alpha = .5, size = 2)+
     geom_smooth(color = "#cc5500", se = FALSE) +
    geom_smooth(color = "#323232", linetype = "dashed", aes(y = landscape_f412), se = FALSE) +
  theme_bw(base_size = 13)+
  #scale_color_manual(values = colors)+ #colors for vegtypes
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  labs(y=expression("BP"[">1.2m"]), x = "") +
  theme(plot.title = element_text(hjust = 0.5))
)

(p_landscapeF8 <- ggplot(data = df_new_landscape, aes(x = focal_inv, y = landscape_f810))+
  #geom_jitter(alpha = .5, size = 2)+
     geom_smooth(color = "#cc5500", se = FALSE) +
    geom_smooth(color = "#323232", linetype = "dashed", aes(y = landscape_f812), se = FALSE) +
  theme_bw(base_size = 13)+
  #scale_color_manual(values = colors)+ #colors for vegtypes
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  labs(y=expression("BP"[">2.4m"]), x = "") +
  theme(plot.title = element_text(hjust = 0.5))
)

```


Combine plots
```{r fig.height=4, fig.width=8}

#create common x axis label
x.grob2 <- textGrob("Proportion of landscape invaded", vjust = -1)


#combine plots
p_alllandscape <- plot_grid(p_landscapeBP, p_landscapeF4, p_landscapeF8 + theme(legend.position = "none"),
                    ncol = 3,
                    nrow = 1,
                    labels = c('(a)', '(b)', '(c)'),
                    label_x = 0, label_y = 0,
                    hjust  = -0.5, vjust = -2.5)

p <- plot_grid(p_alllandscape, nrow = 1, rel_heights = c(1, .2))
```

save plot
```{r}
svg(file = 'Landscapeneighborhood_burn_metrics_ALL.svg', width = 7.8, height = 2.7)

#add common x label
grid.arrange(arrangeGrob(p, bottom = x.grob2))

dev.off()
```




### Extract coef and r2 from gams
```{r}
#combine gams into list


modlist <- list(
             mod_gambpfoa10_landscape, mod_gambpfoa12_landscape,
             mod_gamF4foa10_landscape, mod_gamF4foa12_landscape, 
             mod_gamF8foa10_landscape, mod_gamF8foa12_landscape, 
             #focal forest
            mod_gambpfoa10, mod_gambpfoa12, 
            mod_gamF4foa10, mod_gamF4foa12, 
            mod_gamF8foa10, mod_gamF8foa12,
            #patch
            mod_gamBP10patch,mod_gamBP12patch,
            mod_gamF410patch, mod_gamF412patch)#landscape

modlist_summary = lapply(modlist, summary.Gam)

modnames <- c("landscape BP dif.",  "landscape CBP>1.2m dif.", "landscape CBP>2.4m dif.",
             "invaded landscape BP", "uninvaded landscape BP",
             "invaded landscape CBP>1.2m", "uninvaded landscape CBP>1.2m", 
             "invaded landscape CBP>2.4m", "uninvaded landscape CBP>2.4m", 
            "forest BP dif.", "forest CBP>1.2m dif." , "forest CBP>2.4m dif", #focal forest
            "invaded forest BP", "uninvaded forest BP", 
            "invaded forest CBP>1.2m", "uninvaded forest CBP>1.2m", 
            "invaded forest CBP>2.4m", "uninvaded forest CBP>2.4m",
            "patch BP dif.", "patch CBP>1.2m dif.", #patch
            "invaded patch BP dif.", "uninvaded patch BP dif.",
            "invaded patch CBP>1.2m", "uninvaded patch CBP>1.2m")#landscape

library(texreg)
texreg1 <- htmlreg(modlist, 
                single.row = T,
                custom.model.names = modnames,
                leading.zero = F,
                digits = 3,
                file = 'gams_binom.doc')

summary(mod_gamF410patch)
```


             mod_gambpfoa10_landscape, mod_gambpfoa12_landscape,
             mod_gamF4foa10_landscape, mod_gamF4foa12_landscape, 
             mod_gamF8foa10_landscape, mod_gamF8foa12_landscape, 
            mod_gamBP, mod_gamF4, mod_gamF8, #focal forest
            mod_gambpfoa10, mod_gambpfoa12, 
            mod_gamF4foa10, mod_gamF4foa12, 
            mod_gamF8foa12, mod_gamF8foa10,
            mod_gamBPpatch, mod_gamF4patch, #patch
            mod_gamBP10patch,mod_gamBP12patch,
            mod_gamF410patch, mod_gamF412patch)





---------------------------------------------------------------------------------
### Landscape Focal Analysis with only uninvaded areas

There wasn't much difference in BP when we removed the invaded areas from the analysis. we're not using this analysis in the paper 

mask out invaded pixels from burn metric rasters
```{r}

#mask
BPrastersub <- mask(BPraster, invraster_BME, maskvalue = 1)
plot(BPrastersub)

F4rastersub <- mask(F4raster, invraster_BME, maskvalue = 1)

FLBPrastersub <- mask(FLBPraster, invraster_BME, maskvalue = 1)

```


Average landscape burn metrics for 9x9 pixel landscape (without invaded pixels)

```{r}
focal_BPsub <- raster::focal(BPrastersub, fun = "mean", w=matrix(1,9,9), na.rm = TRUE) 
focal_F4sub <- raster::focal(F4rastersub, fun = "mean", w=matrix(1,9,9), na.rm = TRUE)
focal_FLBPsub <- raster::focal(FLBPrastersub, fun = "mean", w=matrix(1,9,9), na.rm = TRUE)

plot(focal_BPsub)
```

plot relationship between difference in neighborhood burn metrics  and amount of invaded cells within 1000m

extract values from focal raster to spatial dataframe
```{r}
#extract focal burn metrics to the spatial dataframe
spdf$focal_BPsub <- raster::extract(focal_BPsub, sp) 
spdf$focal_F4sub <- raster::extract(focal_F4sub, sp) 
spdf$focal_FLBPsub <- raster::extract(focal_FLBPsub, sp) 


#subset so only areas with > 0 invaded area within neighborhood are included in plot
spdf_sub <- data.frame(spdf[which(spdf$focal_inv > 0),])
#summary(spdf_sub)

```


### Plot neighborhood (without invaded areas)


plot burn metric (absolute difference) response to focal neighborhood - with invaded areas removed
```{r}
p4n <- p_neighborhood2(spdf_sub$focal_BPsub) + ylab("BP (dif.)") + theme(legend.position = "none")

p5n <- p_neighborhood2(spdf_sub$focal_F4sub) + ylab("FLEP>1.2m (dif.)") + theme(legend.position = "none")

p6n <- p_neighborhood2(spdf_sub$focal_FLBPsub) + ylab("joint prob (abs. dif.)") 

```


Combine plots
```{r fig.height=4, fig.width=8}

#create common x axis label
x.grob2 <- textGrob("Proportion of neighborhood invaded", vjust = -4)
legend2 <- get_legend(p6n + theme(legend.position = "bottom", legend.title = element_blank()))

#combine plots
p_all3n <- plot_grid(p4n, p5n, p6n  + theme(legend.position = "none"),
                    ncol = 3,
                    nrow = 1,
                    labels = c('(a)', '(b)', '(c)'),
                    label_x = 0, label_y = 0,
                    hjust  = -0.5, vjust = -1.5)

p <- plot_grid(p_all3n, legend2, nrow = 2, rel_heights = c(1, .1))
```


Not much different for burn probability than when invaded areas are included so did not incldude these analyses in the ms


## Adjacent to invaded areas analysis

buffer invaded areas to evaluate change in areas adjacent to invasion

```{r}
#extract invaded areas
spinv <- spdf %>% subset(alteredFM2 == 1)
spinv <- spinv[,c(2:4, 12)]

#assign projection
crs(spdf) <- CRS("+proj=utm +zone=11 +datum=NAD83 +units=m +no_defs")
crs(spinv) <- CRS("+proj=utm +zone=11 +datum=NAD83 +units=m +no_defs")
# plot(spinv)

#buffer points by 3k
spinv.buf3k <- buffer(spinv, width=3000, dissolve=TRUE)
plot(spinv.buf3k)


#extract points that fall within the 3k buffer
dfextract.invbuf <- raster::intersect(spdf, spinv.buf3k)

#remove invaded pixels
dfinvbuf3k.only <- dfextract.invbuf %>% subset(is.na(alteredFM2))
dfinvbuf3k.only <- data.frame(dfinvbuf3k.only)
#summarize burn metrics for adjacent areas

dfinvbuf3k.only %>% group_by(reclass) %>%
  summarise(n = n(),
            perc = n()/nrow(dfinvbuf3k.only))


dfinvbuf3k.only %>%
  summarize(mean_F4dif = mean(difFILg4), #calculate absolute difference FIL > 4gt
              mean_BPdif = mean(difBP), #calculate absolute difference BP
              mean_F8dif = mean(difFIL8),
              F4_percdif = ((mean(foa10FILg4) - mean(foa12FILg4)) / (mean(foa12FILg4)))*100,
              F8_percdif = ((mean(foa10FIL8) - mean(foa12FIL8)) / (mean(foa12FIL8)))*100,
              BP_percdif = ((mean(foa10BP) - mean(foa12BP)) / (mean(foa12BP)))*100)

```